### 2023-11-22 Scripts/Data_Management/Scripts_Processing/Prep_TyrerCuzick_input.Rmd
Initial commit

---
title: "Prepare data inputs for Tyrer-Cuzick software v8"
author: "Jennifer Collister"
date: "`r Sys.Date()`"
output: 
  word_document:
      fig_width: 8
      fig_height: 5
      reference_docx: "K:/TEU/TEU_Guides/TEU_DocStyle_Rmd_2020.dotx"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include=TRUE, echo=FALSE)

library(kableExtra)
library(knitr)
library(yaml)
library(here)
library(ggplot2)
library(tidyr)
library(readr)
library(dplyr)
library(glue)

# Pretty plots
library(etm)

# Pretty tables
library(arsenal)
library(pander)

# Survival 
library(survival)
library(survminer)

# Flowchart
library(DiagrammeR)
library(DiagrammeRsvg)
library(svglite)
library(rsvg)
library(png)

# Set the root directory to the project directory (otherwise knitr works in the directory of the Rmd file by default)
knitr::opts_knit$set(root.dir = here::here())

# Load the project config file for filepaths etc
config = yaml.load_file(here::here("./config.yml"))
source(here::here(file.path(config$functions)))

#-------------------------------------------------------------------------------------------------------------------

# Output Formatting

# Specify the markdown format for knitr tables, otherwise they're not compatible with kableExtra
options(knitr.kable.NA='', knitr.table.format = "markdown")
options(kableExtra.auto_format = FALSE)

# Pander options
panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)
panderOptions('missing', '')

# Set formatting options for tableby objects
my_controls <- tableby.control(
    test = F,
    total = T,
    #numeric.test = "kwt", cat.test = "chisq",
    numeric.stats = c("meansd", "Nmiss"), #"medianq1q3", 
    cat.stats = c("countpct", "Nmiss"),
    stats.labels = list(
      meansd = "Mean (SD)",
      # medianq1q3 = "Median (IQR)",
      # range = "Min - Max",
      Nmiss = "Missing",
      Npct="N (Pct)"
    ),
    digits = 2L
  )

#-------------------------------------------------------------------------------------------------------
# Setup

figure <- 1
table <- 1
suppfig <- 1
footnote_no <- 1

```

```{r derive-data, include=FALSE, cache=TRUE}

source(file.path(config$scripts$cleaning, "dataset_generator.R"))

exclusions <- function(data) {

  excl$initial <<- nrow(data)

  # Restrict to females (by self-report and genetic)
  data <- data[data$BaC_Sex == "Female",]
  data <- data[data$GeP_Sex == "Female" & !is.na(data$GeP_Sex),]
  excl$sex <<- nrow(data)

  # Restrict to post-menopausal women
  data <- data[!is.na(data$FSF_MenopauseStatus) & data$FSF_MenopauseStatus == "Yes",]
  excl$meno <<- nrow(data)

  # Restrict to White British (genetically)
  # data <- data[!is.na(data$GeP_ethnic),]
  excl$whitebrit <<- nrow(data)

  # Exclude those outside the 40-70 age range
  data <- data[!is.na(data$TEU_BaC_AgeAtRec) & data$TEU_BaC_AgeAtRec >= 40 & data$TEU_BaC_AgeAtRec < 70,]
  excl$agerange <<- nrow(data)
  # Restrict population to those for whom we have PRS scores available
  data <- data[!is.na(data$PRS_bc_std),]
  excl$prs <<- nrow(data)
  # Exclude women with 20+ sisters as the TC model can't handle it
  data <- data[is.na(data$TC_NSisters) | data$TC_NSisters < 20,]
  # Exclude those who didn't answer HRT use
  data <- data[data$TC_HRT_raw != "Unanswered",]
  # Or who reported impossible HRT dates
  data <- data[is.na(data$FSF_HRTAgeStop) | data$FSF_HRTAgeStop <= ceiling(data$TEU_BaC_AgeAtRec),]
  excl$hrt <<- nrow(data)

  # Exclude prior breast cancer
  data <- data[is.na(data$BrCaDx_Prevalent),]
  data <- data[is.na(data$VI_BrCa),]
  excl$brca <<- nrow(data)
  # Prior mastectomy
  data <- data[data$TEU_HES_Mast_prev == "No",]
  excl$mast <<- nrow(data)
  # Prior carcinoma in situ
  data <- data[is.na(data$BrCaInSitu_Prevalent),]
  excl$cainsitu <<- nrow(data)

  return(data)
}

excl <- list(initial=0)

data <-
  evalWithMemoization(
    derive_variables(
      database = config$data$database,
      field_definitions = TEU_SPECS$TyrerCuzick,
      exclusions = exclusions
    ),
    key = c(TEU_SPECS$TyrerCuzick, exclusions)
  )

pretty_names <- pretty_switch(field_definitions=TEU_SPECS$TyrerCuzick)
pretty_list <- pretty_switch(field_definitions=TEU_SPECS$TyrerCuzick, return_type="list")

```

```{r}

table(data$TC_NSisters, useNA='ifany')

tcify <- function(data, nsis) {

  data <- data[data$TC_NSisters == nsis,]

  # Map data to correct format for Tyrer-Cuzick v8 model specification
  tc_data <- data %>% transmute(
    v1_ID = ID,
    v2_Age = floor(replace_na(TEU_BaC_AgeAtRec, -99)),
    v3_AgeMenarche = replace_na(FSF_AgeMenarche, -99),
    v4_Parity = dplyr::case_when(
      TC_Parity == "Nulliparous" ~ 0,
      TC_Parity == "Parous" ~ 1,
      TC_Parity == "Missing" ~ 2),
    v5_AgeFirstBirth = dplyr::case_when(
      TC_Parity == "Nulliparous" ~ 0,
      TC_Parity == "Missing" ~ 0,
      TC_Parity == "Parous" & !is.na(AgeFirstBirth) ~ as.numeric(AgeFirstBirth),
      TC_Parity == "Parous" & is.na(AgeFirstBirth) ~ -99
    ),
    v6_MenoStatus = 2, # we restricted to post-menopausal at baseline
    v7_MenoAge = replace_na(FSF_AgeMenopause, -99), # All women post-menopausal, so no need to handle unknown menopause status
    v8_Height = replace_na(Height, -99),
    v9_Weight = replace_na(Weight, -99),
    v10_Hyperplasia = 0, # missing or unknown
    v11_AtypHyperplasia = dplyr::case_when(
      AtypicalHyperplasia == "Yes" ~ 1,
      TRUE ~ 0
    ),
    v12_LCIS = 0, # missing or unknown - we excluded all CIS at baseline, very few were LCIS
    v13_NOTINUSE = 0, #
    v14_OvarianCancer = dplyr::case_when(
      TC_OvarianCaDx=="Yes" ~ 1,
      TC_OvarianCaDx=="No" ~ 0,
      ),
    v15_OvarianAge = floor(replace_na(TC_OvarianCaAge, -99)),
    v16_Ashkenazi = 0, # missing or unknown
    v17_HRTuse = dplyr::case_when(
      TC_HRT_imputed == "Never" ~ 0,
      TC_HRT_imputed == "Previous user (more than 5 years ago)" ~ 1,
      TC_HRT_imputed == "Previous user (less than 5 years ago)" ~ 2,
      TC_HRT_imputed == "Current" ~ 3
    ),
    v18_HRTType = 1, # missing or unknown
    v19_HRTDuration = round(replace_na(FSF_HRTDuration, 0),1),
    v20_HRTIntended = 0, # missing or unknown
    v21_HRTTimeSince = ifelse(!is.na(FSF_HRT_timesince) & FSF_HRT_timesince < 5,
                              round(FSF_HRT_timesince,1), 0),
    v22_GeneticTest = 0, # unknown

    # Family history
    v23_FatherGenetic = 0, # unknown
    v24_MotherBrCa = dplyr::case_when(
      TC_MotherBrCa=="Yes" ~ 1,
      TC_MotherBrCa=="No" ~ 0,
      ),
    v25_MotherBilateral = 0, # unknown
    v26_MotherOvarian = 0,
    v27_MotherBrCaAge = ifelse(TC_MotherBrCa == "Yes", -99, replace_na(TC_MotherAge, -99)),
    v28_MotherBilAge = -99,
    v29_MotherOvAge = -99,
    v30_MotherGenetic = 0, # unknown

    # Sisters
    v31_NSisters = replace_na(TC_NSisters, 0),
    # This gets complicated now, different number of columns depending on number of sisters
  )

  # This part is different depending on how many sisters
  sisters <- data %>% transmute(
    v32_SisterBrCa = dplyr::case_when(
      TC_SiblingBrCa=="Yes" ~ 1,
      TC_SiblingBrCa=="No" ~ 0,
      ),
    v33_SisterBilateral = 0,
    v34_SisterOvarian = 0,
    v35_SisterBrCaAge = -99,
    v36_SisterBilAge = -99,
    v37_SisterOvAge = -99,
    v38_SisterGenetic = 0
  )

  if(nsis == 1) {
    tc_data <- cbind(tc_data, sisters)
  } else if (nsis > 1) {
    extra <- sisters %>% mutate(v32_SisterBrCa = 0)
    for (n in c(1: (nsis-1))) {
      colnames(extra) <- str_replace(colnames(extra),
                                     "[0-9]+",
                                     function(x) {return(as.numeric(x) + 7)}
                                     )
      sisters <- cbind(sisters, extra)
    }
    tc_data <- cbind(tc_data, sisters)
  }

  # And the rest of the compulsory family history
  morefam <- data %>% transmute(
    # Paternal Gran
    s2_PatGranBrCa = 0,
    s3_PatGranBilateral = 0,
    s4_PatGranOvarian = 0,
    s5_PatGranBrCaAge = -99,
    s6_PatGranBilAge = -99,
    s7_PatGranOvAge = -99,
    s8_PatGranGenetic = 0,

    # Maternal gran
    s9_MatGranBrCa = 0,
    s10_MatGranBilateral = 0,
    s11_MatGranOvarian = 0,
    s12_MatGranBrCaAge = -99,
    s13_MatGranBilAge = -99,
    s14_MatGranOvAge = -99,
    s15_MatGranGenetic = 0,

    # Aunts
    s16_NPatAunts = 0,
    # s17_PatAuntBrCa = 0,
    # s18_PatAuntBilateral = 0,
    # s19_PatAuntOvarian = 0,
    # s20_PatAuntBrCaAge = -99,
    # s21_PatAuntBilAge = -99,
    # s22_PatAuntOvAge = -99,
    # s23_PatAuntGenetic = 0,

    # Maternal Aunts
    pa1_NMatAunts = 0,
    # pa2_MatAuntBrCa = 0,
    # pa3_MatAuntBilateral = 0,
    # pa4_MatAuntOvarian = 0,
    # pa5_MatAuntBrCaAge = -99,
    # pa6_MatAuntBilAge = -99,
    # pa7_MatAuntOvAge = -99,
    # pa8_MatAuntGenetic = 0,

    # Daughters
    ma1_NDaughters = 0,
    # ma2_DaughterBrCa = 0,
    # ma3_DaughterBilateral = 0,
    # ma4_DaughterOvarian = 0,
    # ma5_DaughterBrCaAge = -99,
    # ma6_DaughterBilAge = -99,
    # ma7_DaughterOvAge = -99,
    # ma8_DaughterGenetic = 0
  )
  tc_data <- cbind(tc_data, morefam)

  table3 <- data %>% transmute(
    zero1 = nsis,
    zero2 = 0,
    zero3 = 0,
    zero4 = 0,
    zero5 = 0,
    zero6 = 0,
    zero7 = 0
  ) # No participants have the optional family history data. Spec says 7 columns of 0
  if (nsis > 0) {
    for (sis in c(1:nsis)) {
      table3[paste0("sis", sis)] <- 0 
    }
  } # Technically these columns should come after zero1 but they're all just 0 so it doesn't matter
  tc_data <- cbind(tc_data, table3)

  v8 <- data %>% transmute(
    t1_FatherBrCa = dplyr::case_when(
      TC_FatherBrCa=="Yes" ~ 1,
      TC_FatherBrCa=="No" ~ 0,
      ), # All No but include anyway for completion
    t2_BrotherBrCa = dplyr::case_when(
      TC_SiblingBrCa == "Yes" & 
        (TC_NSisters == 0 | is.na(TC_NSisters)) & 
        (TC_NBrothers > 0 & !is.na(TC_NBrothers)) ~ 1,
      TRUE ~ 0
    ), # If siblings with breast cancer, no sisters, and brothers then yes, otherwise no
    t3_DensityMeasure = 0,
    t4_DensityValue = -99,
    t5_PRS = 0 #round(TEU_BrCa_313_PRS_center,8)
  )

  tc_data <- cbind(tc_data, v8)
  return(tc_data)
}

nrows <- nrow(data)
chunksize <- 15000
start <- 1
ordered <- data %>% # Treat NA sisters as 0
  mutate(TC_NSisters = ifelse(is.na(TC_NSisters), 0, TC_NSisters)) %>%
  arrange(TC_NSisters)

while(start < nrows) {

  end <- min(start + chunksize - 1, nrows)
  filename <- file.path(config$data$derived, "noPRS", paste0("rows", start, "_", end, ".txt"))

  chunk <- ordered[start:end,]
  nsisters <- sort(unique(chunk$TC_NSisters))
  nparticipants <- nrow(chunk)

  header <- data.frame(col1 = c("v8", nparticipants))
  write.table(header, file=filename, append=FALSE,
              sep="\t", eol="\n", quote=FALSE,
              row.names=FALSE, col.names=FALSE)

  for (n in nsisters) {
    output <- tcify(chunk, nsis=n)

    write.table(output, file=filename, append=TRUE,
              sep="\t", eol="\n", quote=FALSE,
              row.names=FALSE, col.names=FALSE)
  }
  start <- end + 1
}

#-------------#
# Run 3 times #
#-------------#
# t5_PRS = 0, filename "noPRS"
# t5_PRS = TEU_BrCa_313_PRS_center, filename "withPRS/Mavaddat2019"
# t5_PRS = TEU_BrCa_100k_PRS_center, filename = "withPRS/Fritsche2020"
# Note that the different PRS are on different scales

```

```{r}
# Read Tyrer-Cuzick risks output file

filenames <- list.files(file.path(config$data$derived, "noPRS", "risk"), 
                        full.names=TRUE, recursive=TRUE, include.dirs=FALSE)

allrisks <- lapply(filenames, function(x){
  read.table(x, header=FALSE, sep="\t",
             col.names=c("ID", "nyear", "ltime_risk", "pop_ltime_risk", "brca1", "brca2"))
  })

risks <- dplyr::bind_rows(allrisks)

```