### 2024-02-10 Scripts/Stats_Analysis/RMarkdown/ManuscriptOutputs.Rmd
Tidy figures, save in format required for resubmission

---
title: "Manuscript_Outputs"
author: "Jennifer Collister"
date: "`r Sys.Date()`"
output:
    word_document:
      fig_width: 8
      fig_height: 5
      reference_docx: "K:/TEU/TEU_Guides/TEU_DocStyle_Rmd_2020.dotx"
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(include=TRUE, echo=FALSE, message = FALSE)

library(kableExtra)
# Issue with kableExtra collapse_rows for data dictionary
# Use development version
# renv::install("haozhu233/kableExtra@a6af5c0")
# See https://github.com/haozhu233/kableExtra/issues/595#issuecomment-808703159

library(knitr)
library(yaml)
library(here)
library(ggplot2)
library(tidyr)
library(readr)
library(dplyr)
library(glue)
library(tictoc)

# Pretty plots
library(etm)
library(plotROC) # For XML dependency, used install.packages("XML", type = "binary") to avoid compilation issues
library(gridExtra)
library(maps)
library(timeROC)

# Pretty tables
library(arsenal)
library(pander)
library(flextable)

# Survival 
library(survival)
library(survminer)
library(rms) # For Hmisc dependency, used install.packages("Hmisc@4.7-2") to avoid installation issues
library(lmtest)

# NRI
library(nricens)

# Gail model
library(BCRA)

# Flowchart
library(DiagrammeR)
library(DiagrammeRsvg)
library(svglite)
library(rsvg)
library(png)

# Set the root directory to the project directory (otherwise knitr works in the directory of the Rmd file by default)
knitr::opts_knit$set(root.dir = here::here())

# Load the project config file for filepaths etc
config = yaml.load_file(here::here("./config.yml"))
source(here::here(file.path(config$functions)))

#-------------------------------------------------------------------------------------------------------------------

# Output Formatting

# Specify the markdown format for knitr tables, otherwise they're not compatible with kableExtra
options(knitr.kable.NA='', knitr.table.format = "markdown")
options(kableExtra.auto_format = FALSE)

# Pander options
panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)
panderOptions('missing', '')

# Set formatting options for tableby objects
my_controls <- tableby.control(
    test = F,
    total = T,
    #numeric.test = "kwt", cat.test = "chisq",
    numeric.stats = c("meansd", "Nmiss"), #"medianq1q3", 
    cat.stats = c("countpct", "Nmiss"),
    stats.labels = list(
      meansd = "Mean (SD)",
      # medianq1q3 = "Median (IQR)",
      # range = "Min - Max",
      Nmiss = "Missing",
      Npct="N (Pct)"
    ),
    digits = 2L
  )

by_status <- tableby.control(
    test = T,
    total = T,
    numeric.test = "kwt", cat.test = "chisq",
    numeric.stats = c("meansd", "Nmiss"), #"medianq1q3", 
    cat.stats = c("countpct", "Nmiss"),
    stats.labels = list(
      meansd = "Mean (SD)",
      # medianq1q3 = "Median (IQR)",
      # range = "Min - Max",
      Nmiss = "Missing",
      Npct="N (Pct)"
    ),
    digits = 2L
  )

#-------------------------------------------------------------------------------------------------------
# Setup

figure <- 1
table <- 1
suppfig <- 1
footnote_no <- 1

```

```{r derive-data, include=FALSE}

source(file.path(config$scripts$cleaning, "dataset_generator.R"))

exclusions <- function(data) {

  excl$initial <<- nrow(data)

  # Restrict to females (by self-report and genetic)
  data <- data[data$BaC_Sex == "Female",]
  data <- data[data$GeP_Sex == "Female" & !is.na(data$GeP_Sex),]
  excl$sex <<- nrow(data)

  # Restrict to post-menopausal women
  data <- data[!is.na(data$FSF_MenopauseStatus) & data$FSF_MenopauseStatus == "Yes",]
  excl$meno <<- nrow(data)

  # Restrict to White British (genetically)
  data <- data[!is.na(data$GeP_ethnic),]
  excl$whitebrit <<- nrow(data)

  # Exclude those outside the 40-70 age range
  data <- data[!is.na(data$TEU_BaC_AgeAtRec) & data$TEU_BaC_AgeAtRec >= 40 & data$TEU_BaC_AgeAtRec < 70,]
  excl$agerange <<- nrow(data)
  # Restrict population to those for whom we have PRS scores available
  data <- data[!is.na(data$PRS_bc_std),]
  excl$prs <<- nrow(data)
  # Exclude women with 20+ sisters as the TC model can't handle it
  data <- data[is.na(data$TC_NSisters) | data$TC_NSisters < 20,]
  # Exclude those who didn't answer HRT use
  data <- data[data$TC_HRT_raw != "Unanswered",]
  # Or who reported impossible HRT dates
  data <- data[is.na(data$FSF_HRTAgeStop) | data$FSF_HRTAgeStop <= ceiling(data$TEU_BaC_AgeAtRec),]
  excl$hrt <<- nrow(data)

  # Exclude prior breast cancer
  data <- data[is.na(data$BrCaDx_Prevalent),]
  data <- data[is.na(data$VI_BrCa),]
  excl$brca <<- nrow(data)
  # Prior mastectomy
  data <- data[data$TEU_HES_Mast_prev == "No",]
  excl$mast <<- nrow(data)
  # Prior carcinoma in situ
  data <- data[is.na(data$BrCaInSitu_Prevalent),]
  excl$cainsitu <<- nrow(data)

  return(data)
}

excl <- list(initial=0)
tic("Data derived: ")
data <- derive_variables(
  database = config$data$database,
  field_definitions = TEU_SPECS$PRSforBrCa,
  exclusions = exclusions,
  dictionary = "PRSforBrCa.html"
)
toc()

pretty_names <- pretty_switch(field_definitions=TEU_SPECS$PRSforBrCa)
pretty_list <- pretty_switch(field_definitions=TEU_SPECS$PRSforBrCa, return_type="list")

# Standardise variables so HR are per sd
data <- data %>% mutate(
  TC_10risk_sd = TC_10risk/sd(TC_10risk),
  gail_10risk_sd = gail_10risk/sd(gail_10risk),
  PRS_bc_sd = PRS_bc_std/sd(PRS_bc_std),
  TEU_BrCa_313_PRS_center = TEU_BrCa_313_PRS_center / sd(TEU_BrCa_313_PRS_center),
  TEU_BrCa_100k_PRS_center = TEU_BrCa_100k_PRS_center / sd(TEU_BrCa_100k_PRS_center)
  )

# Separate out test and training data
# train_IDs <- data$ID[data$train_set==TRUE]
# write.csv(train_IDs, file.path(config$data$derived, "training_ids.csv"), row.names = FALSE)
train_IDs <- read.csv(file.path(config$data$derived, "training_ids.csv"))$x

train_data <- data %>% filter(ID %in% train_IDs)

test_data <- data %>% filter(!ID %in% train_IDs)

```

```{r}
label1 <- paste0('UKB participants\\nn = ', excl$initial)
label2 <- paste0('Post-menopausal\nWhite British\nfemale participants\\nn = ', excl$agerange)
label3 <- paste0('Aged between 40 and 70\nwith available core data\\nn = ', excl$hrt)
label4 <- paste0('Study population: post-menopausal women\nwith no history of breast cancer\\nn = ', excl$cainsitu)

label5 <- paste0(excl$initial-excl$sex, ' males or genetically sex discordant\\l', excl$sex-excl$meno, ' pre-menopausal\\lor unknown menopausal status \\l', excl$meno-excl$whitebrit, ' not of genetically White British ancestry\\l')
label6 <- paste0(excl$whitebrit-excl$agerange,' outside the age range of [40,70) years \\l', excl$agerange-excl$prs,' Genomics plc PRS unavailable\\l', excl$prs-excl$hrt, ' invalid or missing data for core variables*\\l')
label7 <- paste0(excl$hrt-excl$brca, ' previous breast cancer\\l', excl$brca-excl$mast, ' previous mastectomy\\l', excl$mast-excl$cainsitu,' previous carcinoma in situ\\l')

gv <- read_file(here::here(file.path(config$outputs$flowcharts, "Exclusions.gv")))

# Supplementary Figure 1

export_svg(DiagrammeR::grViz(glue(gv))) %>% 
  charToRaw %>% 
  rsvg %>%
  png::writePNG(here::here(file.path(config$outputs$figures$final, "SuppFigureS1.png")))

```

## Cox model containing breast cancer risk score and PRS in training data (N=`r nrow(train_data)`)

```{r, fig.height=6}
# Incorporate PRS using a Cox model in training data, then predict 10-yr risks in the test data
plot_varnames <- function(x) {
  y_names <- dplyr::case_when(
    x == "TC_10risk_sd" ~ "Tyrer-Cuzick 10yr risk",
    x == "gail_10risk_sd" ~ "Gail 10yr risk",
    x == "PRS_bc_sd" ~ "PRS_BC",
    TRUE ~ x
  )
  return(y_names)
}

# Mimic get.risk.coxph from nricens but modify to use prediction in new data
get.risk.coxph_test <- function(mdl, t0, newdata) {
  bash <- basehaz(mdl)
  lambda0 <- approx(bash$time, bash$hazard, t0)$y
  linpred <- predict(mdl, newdata=newdata, type="lp")
  risk <- 1 - exp(-lambda0 * exp(linpred))
}

# Tyrer-Cuzick
model_TC <- coxph(Surv(TEU_BrCa_time, TEU_BrCa_status) ~
                        TC_10risk_sd + PRS_bc_sd + GeP_Array + 
                        GPLC_PC_1 + GPLC_PC_2 + GPLC_PC_3 + GPLC_PC_4,
                 data=train_data)

test_data$TC_prs_10risk <- get.risk.coxph_test(model_TC, t0=10, newdata=test_data)
# train_data$TC_prs_10risk2 <- predict(model_TC, type = "risk")

# plot Schoenfeld residuals
test.ph <- cox.zph(model_TC)
# test.ph
y <- test.ph$y
colnames(y) <- plot_varnames(colnames(y))
test.ph$y <- y

# Supplementary Figure 3
tiff(file.path(config$outputs$figures$final,"SuppFigureS3.tiff"), units="in", width=8, height=5, res=300)
ggcoxzph(test.ph, point.alpha=0.04, var=c("Tyrer-Cuzick 10yr risk", "PRS_BC"),
         caption="Model also includes genetic array and first 4 principal components of genetic ancestry, plots omitted.")
dev.off()

# Gail
model_gail <- coxph(Surv(TEU_BrCa_time, TEU_BrCa_status) ~
                        gail_10risk_sd + PRS_bc_sd + GeP_Array + 
                        GPLC_PC_1 + GPLC_PC_2 + GPLC_PC_3 + GPLC_PC_4,
                 data=train_data)

test_data$gail_prs_10risk <- get.risk.coxph_test(model_gail, t0=10, newdata=test_data)

# plot Schoenfeld residuals
test.ph <- cox.zph(model_gail)
# test.ph
y <- test.ph$y
colnames(y) <- plot_varnames(colnames(y))
test.ph$y <- y

# Supplementary Figure 4
tiff(file.path(config$outputs$figures$final,"SuppFigureS4.tiff"), units="in", width=8, height=5, res=300)
ggcoxzph(test.ph, point.alpha=0.04, var=c("Gail 10yr risk", "PRS_BC"),
         caption="Model also includes genetic array and first 4 principal components of genetic ancestry, plots omitted.")
dev.off()

# Test data

test_data <- test_data %>% 
  mutate(
    TC_10risk_cat = factor(ifelse(TC_10risk_cal <= 0.05, "<=5%", ">5%"),
                         levels=c("<=5%", ">5%")),
    gail_10risk_cat = factor(ifelse(gail_10risk_cal <= 0.05, "<=5%", ">5%"),
                           levels=c("<=5%", ">5%")),
    TC_prs_10risk_cat = factor(ifelse(TC_prs_10risk <= 0.05, "<=5%", ">5%"),
                             levels=c("<=5%", ">5%")),
  gail_prs_10risk_cat = factor(ifelse(gail_prs_10risk <= 0.05, "<=5%", ">5%"),
                               levels=c("<=5%", ">5%"))
  )

# Table
get_HR_perSD <- function(model, var) {

  coeff <- summary(model)$coefficients
  HR <- coeff[rownames(coeff)==var,]

  return(paste0(pretty_dp(exp(HR[1]), dp=2), " ",
                pretty_confint(exp(HR[1] - (1.96*HR[3])), 
                               exp(HR[1] + (1.96*HR[3])), dp=2))
  )
}

dt <- data.frame(var=c("Tyrer-Cuzick", "Gail"),
                 model=c(get_HR_perSD(model_TC, "TC_10risk_sd"), get_HR_perSD(model_gail, "gail_10risk_sd")),
                 prs=c(get_HR_perSD(model_TC, "PRS_bc_sd"), get_HR_perSD(model_gail, "PRS_bc_sd"))
)

ft <- flextable(dt) %>%
  set_header_labels(var="Model", model="10-year risk score HR per SD", prs="PRS HR per SD") %>%
  set_table_properties(layout="autofit") %>%
  flextable::footnote(i=1, j=1, 
           value=as_paragraph("Hazard ratio per standard deviation from a Cox model containing 10-year risk score from breast cancer risk prediction model, Genomics plc PRS, genetic array and first 4 principal components of genetic ancestry, in the training data"), 
           ref_symbols="a", part="header")

ft

```

```{r}
# How many women are above 5% 10yr risk according to the original model?

TC_5pct <- table(test_data$TC_10risk_cal > 0.05, useNA='ifany')
TC_5prop <- TC_5pct[2]/nrow(test_data)
TC_prs_threshold <- quantile(test_data$TC_prs_10risk, probs=c(1-TC_5prop))

gail_5pct <- table(test_data$gail_10risk_cal > 0.05, useNA='ifany')
gail_5prop <- gail_5pct[2]/nrow(test_data)
gail_prs_threshold <- quantile(test_data$gail_prs_10risk, probs=c(1-gail_5prop))

# How many women are above 5% 10yr risk after adding the PRS?

TC_combined_5pct <- table(test_data$TC_prs_10risk > 0.05)

gail_combined_5pct <- table(test_data$gail_prs_10risk > 0.05)

test_data <- test_data %>% mutate(
  TC_prs_10risk_adj = TC_prs_10risk - (quantile(TC_prs_10risk, probs=c(1-TC_5prop)) - 0.05),
  TC_prs_10risk_adj_cat = factor(ifelse(TC_prs_10risk <= TC_prs_threshold, 
                                          paste0("<=", round(100*TC_prs_threshold,2), "%"), 
                                          paste0(">", round(100*TC_prs_threshold,2), "%")),
                               levels=c(paste0("<=", round(100*TC_prs_threshold,2), "%"), 
                                        paste0(">", round(100*TC_prs_threshold,2), "%"))),
  gail_prs_10risk_adj = gail_prs_10risk - (quantile(gail_prs_10risk, probs=c(1-gail_5prop)) - 0.05),
  gail_prs_10risk_adj_cat = factor(ifelse(gail_prs_10risk <= gail_prs_threshold, 
                                          paste0("<=", round(100*gail_prs_threshold,2), "%"), 
                                          paste0(">", round(100*gail_prs_threshold,2), "%")),
                               levels=c(paste0("<=", round(100*gail_prs_threshold,2), "%"), 
                                        paste0(">", round(100*gail_prs_threshold,2), "%")))
  )

```

```{r, results=FALSE}

# Net reclassification index, for Table 1

# Tyrer-Cuzick
nri_tc <- nricens(time=test_data$TEU_BrCa_time, event=test_data$TEU_BrCa_status,
        p.std=test_data$TC_10risk_cal, p.new=test_data$TC_prs_10risk_adj,
        t0=10, cut=0.05, msg=FALSE)

# Gail
nri_gail <- nricens(time=test_data$TEU_BrCa_time, event=test_data$TEU_BrCa_status,
        p.std=test_data$gail_10risk_cal, p.new=test_data$gail_prs_10risk_adj,
        t0=10, cut=0.05, msg=FALSE)

nri_ci <- function(nri_output, type="overall", dp=3){
  row <- dplyr::case_when(
    type=="overall" ~ 1,
    type=="case" ~ 2,
    type=="control" ~ 3
  )
  est <- nri_output$nri$Estimate[row]
  lci <- nri_output$nri$Lower[row]
  uci <- nri_output$nri$Upper[row]

  return(paste0(pretty_dp(est, dp=dp), " ", pretty_confint(lci, uci, dp=dp)))
}

```

## Table `r table`: Model performance with/without PRS in test data (N=`r nrow(test_data)`)
`r table <- table + 1`

```{r}

dt <- data.frame(model=c("Tyrer-Cuzick", "Model only", "Model with PRS", 
                         "Gail", "Model only", "Model with PRS"),
                 HarrellC=c(NA,
                       HarrellC_CI(df=test_data, score="TC_10risk_cal", 
                                   time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2),
                       HarrellC_CI(df=test_data, score="TC_prs_10risk", 
                                   time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2),
                       NA,
                       HarrellC_CI(df=test_data, score="gail_10risk_cal", 
                                   time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2),
                       HarrellC_CI(df=test_data, score="gail_prs_10risk", 
                                   time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2)),
                 NRI_overall=c(NA, nri_ci(nri_tc, type="overall", dp=3), nri_ci(nri_tc, type="overall", dp=3), 
                               NA, nri_ci(nri_gail, type="overall", dp=3), nri_ci(nri_gail, type="overall", dp=3)),
                 NRI_case=c(NA, nri_ci(nri_tc, type="case", dp=3), nri_ci(nri_tc, type="case", dp=3), 
                               NA, nri_ci(nri_gail, type="case", dp=3), nri_ci(nri_gail, type="case", dp=3)),
                 NRI_control=c(NA, nri_ci(nri_tc, type="control", dp=3), nri_ci(nri_tc, type="control", dp=3), 
                               NA, nri_ci(nri_gail, type="control", dp=3), nri_ci(nri_gail, type="control", dp=3)))

ft <- flextable(dt) %>%
  merge_v(j = ~ model + NRI_overall + NRI_case + NRI_control) %>%
  # separate_header(sep="_") %>%
  set_header_labels(model="Model", HarrellC="Harrell's C (95% CI)",
                    NRI_overall="Overall", 
                    NRI_case=paste0("Case (N=", nrow(test_data[test_data$TEU_BrCa_time <= 10 &
                                                                 test_data$TEU_BrCa_status==1,]), ")"), 
                    NRI_control=paste0("Control (N=", nrow(test_data[test_data$TEU_BrCa_time > 10,]), ")")) %>%
  add_header_row(values=c("", "NRI"), colwidths=c(2, 3)) %>%
  align(align = "center", part = "all", j=c("HarrellC", 'NRI_overall', 'NRI_case', 'NRI_control')) %>%
  bg(i= ~ is.na(HarrellC), bg="lightgrey", part="body") %>%
  merge_at(i=1, j=1:5) %>%
  merge_at(i=4, j=1:5) %>%
  fix_border_issues(part="all") %>%
  set_table_properties(layout="autofit") %>%
  flextable::footnote(i=1, j=3, 
           value=as_paragraph("NRI: Net Reclassification Index for 10yr risks from model with PRS compared to Calibrated model, with a fixed proportion of women classified as high risk"), 
           ref_symbols="a", part="header") %>%
  flextable::footnote(i=2, j=3, 
           value=as_paragraph("Overall NRI = Case NRI + Control NRI, where cases are defined as individuals diagnosed with breast cancer within 10 years and controls are defined as individuals who were still at risk of breast cancer by 10 years of follow-up."), 
           ref_symbols="b", part="header")

ft

# Flextable: Border issues with merged cells
# https://davidgohel.github.io/flextable/reference/fix_border_issues.html

```

## Figure `r figure`: Calibration plots
`r figure <- figure + 1`

```{r, fig.height=7}

# Calibration - code adapted from Ruthie's Project and Xiaonan's diabetes work with Leicester score

cal_plot <- function(data, var="TC_10risk", title, xlim=c(0,0.1), ylim=c(0,0.1)){

  # First group into deciles of predicted risk
  deciles <- quantile(data[[var]], probs = seq(0, 1, 0.1))
  labels <- c("1st decile", "2nd", "3rd", "4th", "5th", "6th", "7th", "8th", "9th", "10th decile")
  data <- data %>% mutate(
    TEU_BrCa_status = ifelse(TEU_BrCa_time < 10, TEU_BrCa_status, 0),
    TEU_BrCa_time = ifelse(TEU_BrCa_time < 10, TEU_BrCa_time, 10),
    risk_dec = cut(get(var), breaks=deciles, labels=labels, right=TRUE, include.lowest=TRUE)
    )

  # Mean predicted risk per decile
  predicted <- data %>%
    group_by(risk_dec) %>% 
    summarise(predicted = mean(get(var)))

  # Fit survival curves from Kaplan-Meier and extract observed risks by predicted risk decile
  fit <- survfit(Surv(TEU_BrCa_time, TEU_BrCa_status) ~ risk_dec, data=data)
  fit_time <- summary(fit, times = 10)

  observed <- data.frame(risk_dec = fit_time$strata,
                         observed = 1 - fit_time$surv,
                         observed_lci = 1 - fit_time$lower,
                         observed_uci = 1 - fit_time$upper) %>%
    mutate(risk_dec = as.factor(str_remove(risk_dec, "risk_dec=")))

  # Join predicted and observed
  calibration <- predicted %>% inner_join(observed, by="risk_dec")

  # Calculate calibration slope
  cal_model <- lm(observed ~ predicted, data=calibration)
  cal_slope <- cal_model$coefficients[2]
  cal_intercept <- cal_model$coefficients[1]

  # Plot in calibration plot
  plot <- ggplot(calibration, aes(predicted, observed)) +
    geom_point(shape=22, size=3, color="blue", fill="blue") +
    geom_abline(aes(slope=1, intercept=0, linetype="dashed", colour="Ideal (y=x)"), show.legend=FALSE) +
    geom_smooth(method=lm, aes(colour="Best fit")) +
    geom_pointrange(aes(ymin=observed_lci, ymax=observed_uci)) +
    scale_x_continuous(limits=xlim) +
    scale_y_continuous(limits=ylim) +
    labs(x="Predicted risk",
         y="Observed risk",
         title=title) +
    annotate("text", x=0.07, y = 0.015, size=10/.pt,
             label=paste0("Slope = ", pretty_dp(cal_slope, 3), 
                          "\nIntercept = ", pretty_dp(cal_intercept, 3))) +
    scale_color_manual(name = "Legend",             # legend name
                     values = c("Ideal (y=x)" = "black",  # map regression line colors
                                "Best fit" = "blue")) +
    theme(legend.justification = c(0, 1), legend.position = c(0, 1)) +
    theme(plot.title = element_text(size=10))

  ggsave(file.path(config$outputs$figures$interim, paste0(var, "_calibration_plot10.jpg")), 
       scale=0.6, width=25, height=15, units="cm")

  output <- list(plot=plot, cal_slope=cal_slope, cal_intercept=cal_intercept)

  return(output)

}

TC_train <- cal_plot(data=train_data, var="TC_10risk",
                   title="Tyrer-Cuzick model in training data")
gail_train <- cal_plot(data=train_data, var="gail_10risk",
                     title="Gail model in training data")

TC_test <- cal_plot(data=test_data, var="TC_10risk_cal",
                   title="Calibrated Tyrer-Cuzick model in test data")
gail_test <- cal_plot(data=test_data, var="gail_10risk_cal",
                     title="Calibrated Gail model in test data")

TC_prs_cal <- cal_plot(data=test_data, var="TC_prs_10risk",
                   title="Tyrer-Cuzick model with PRS in test data")
gail_prs_cal <- cal_plot(data=test_data, var="gail_prs_10risk",
                     title="Gail model with PRS in test data")

# 2. Save the legend
#+++++++++++++++++++++++
legend <- get_legend(TC_train$plot)
# 3. Remove the legend
#+++++++++++++++++++++++
# Figure 1
tiff(file.path(config$outputs$figures$final,"Figure1.tiff"), units="in", width=8, height=10, res=300)
gridExtra::grid.arrange(TC_train$plot + theme(legend.position="none"), 
                        gail_train$plot + theme(legend.position="none"), 
                        legend, 
                        TC_test$plot + theme(legend.position="none"), 
                        gail_test$plot + theme(legend.position="none"),
                        NULL,
                        TC_prs_cal$plot + theme(legend.position="none"),
                        gail_prs_cal$plot + theme(legend.position="none"),
                        ncol=3, widths=c(3.5, 3.5, 1))
dev.off()

```

## Figure `r figure`: ROC Plots
`r figure <- figure + 1`

```{r, fig.width=12, warning=FALSE}

# Time-dependent ROC at 10 years, using timeROC

td_roc_plot <- function(data, modelname, prsname, plotname,
                     varlist=c("gail_10risk_cal", "PRS_bc_std", "gail_prs_10risk")){

  # Time-dependent ROC
  tROC <- lapply(varlist,
                 function(i) timeROC::timeROC(T=data$TEU_BrCa_time, 
                                              delta=data$TEU_BrCa_status, 
                                              marker=data[[i]],
                                              cause=1, times=c(9, 10), weighting="marginal"))

  tROC_long <- rbind(cbind(tROC[[1]][["TP"]], tROC[[1]][["FP"]]),
                     cbind(tROC[[2]][["TP"]], tROC[[2]][["FP"]]),
                     cbind(tROC[[3]][["TP"]], tROC[[3]][["FP"]]))
  colnames(tROC_long) <- c("TP_9", "TP_10", "FP_9", "FP_10")
  tROC_df <- data.frame(tROC_long)
  tROC_df$Model <- factor(c(rep(paste0(modelname, " 10-yr risk"), times=dim(tROC[[1]][["TP"]])[1]),
                     rep(prsname, times=dim(tROC[[2]][["TP"]])[1]),
                     rep(paste0(modelname, " & PRS_BC 10-yr risk"), times=dim(tROC[[3]][["TP"]])[1])),
                     levels=c(paste0(modelname, " 10-yr risk"),
                              prsname,
                              paste0(modelname, " & PRS_BC 10-yr risk")))

  plot <- ggplot(tROC_df,aes(x=FP_10, y=TP_10, color=Model))+
    geom_line(linewidth=1) + style_roc() +
    geom_abline(aes(slope=1, intercept=0), linetype="dashed") +
    xlab("False positive fraction") + ylab("True positive fraction") +
    scale_color_manual(name="Model",values = c("#F8766D","#00BA38", "#6193FF")) +
    labs(title =plotname, caption = "Dashed line is y=x")+
    theme(legend.position = c(0.7, 0.15))

  ggsave(file.path(config$outputs$figures$interim, paste0("rocplot_", modelname, ".jpg")),
       scale=0.7, width=25, height=15, units="cm")

  plot
}

# Tyrer-Cuzick
TC_roc <- td_roc_plot(test_data, modelname="Tyrer-Cuzick", prsname="PRS_BC", plotname="Tyrer-Cuzick",
         varlist=c("TC_10risk_cal", "PRS_bc_std", "TC_prs_10risk"))

# Gail
gail_roc <- td_roc_plot(test_data, modelname="Gail", prsname="PRS_BC", plotname="Gail",
         varlist=c("gail_10risk_cal", "PRS_bc_std", "gail_prs_10risk"))

# Figure 2
tiff(file.path(config$outputs$figures$final,"Figure2.tiff"), units="in", width=10, height=5, res=300)
gridExtra::grid.arrange(TC_roc, gail_roc,
                        ncol=2)
dev.off()

```

## Tables 3 Reclassification

### Tyrer-Cuzick

When using the Tyrer-Cuzick model, `r pretty_dp(100*TC_5prop, 2)`% of women are above 5% 10yr risk.

This rises to `r pretty_dp(100*TC_combined_5pct[2]/nrow(test_data),2)`% of women above this 5% threshold when using the combination of both Tyrer-Cuzick and PRS in a Cox model.

Instead of using a fixed threshold, let's use the same proportion of women.

NRI using top `r pretty_dp(100*TC_prs_threshold,2)`% of 10yr risk as risk threshold

```{r}

nri_tab <- function(p, population, model_old, model_new) {
  p <- propped(p)
  p <- as.data.frame(cbind(c(model_old, model_old, ""), rownames(p), p))

  flextable(p) %>%
    merge_v(j = ~ V1) %>%
    set_header_labels(V1="", V2="") %>%
    add_header_row(values=c(population, model_new, ""), colwidths=c(2, 2,1)) %>%
    align(align = "center", part = "all") %>%
    set_table_properties(layout="autofit") %>%
    border_remove() %>%
    hline_top() %>%
    hline(i=c(2)) %>%
    vline(j=c(2, 4)) %>%
    fix_border_issues(part="all") %>%
    flextable::footnote(i=1, j=1, value=as_paragraph("Cases defined as individuals diagnosed with breast cancer within 10 years. Controls defined as individuals who were still at risk of breast cancer by 10 years of follow-up. Individuals censored before 10 years are not displayed."), 
           ref_symbols="a", part="header")
}

################
# Tyrer-Cuzick #
################

# Cases
print("Cases")
nri_tab(table(test_data$TC_10risk_cat[test_data$TEU_BrCa_time <= 10 & test_data$TEU_BrCa_status==1], 
              test_data$TC_prs_10risk_adj_cat[test_data$TEU_BrCa_time <= 10 & test_data$TEU_BrCa_status==1]), 
        population="Cases", model_old="Tyrer-Cuzick", model_new="Tyrer-Cuzick + PRS")

# Controls
print("Controls")
nri_tab(table(test_data$TC_10risk_cat[test_data$TEU_BrCa_time > 10], 
              test_data$TC_prs_10risk_adj_cat[test_data$TEU_BrCa_time > 10]), 
        population="Controls", model_old="Tyrer-Cuzick", model_new="Tyrer-Cuzick + PRS")

```

### Gail

When using the Gail model, `r round(100*gail_5prop, 2)`% of women are above 5% 10yr risk.

This rises to `r round(100*gail_combined_5pct[2]/nrow(test_data),2)`% of women above this 5% threshold when using the combination of both Gail and PRS in a Cox model.

Instead of using a fixed threshold, let's use the same proportion of women.

NRI using top `r round(100*gail_prs_threshold,2)`% of 10yr risk as risk threshold

```{r}

########
# Gail #
########

# Cases
print("Cases")
nri_tab(table(test_data$gail_10risk_cat[test_data$TEU_BrCa_time <= 10 & test_data$TEU_BrCa_status==1], 
              test_data$gail_prs_10risk_adj_cat[test_data$TEU_BrCa_time <= 10 & test_data$TEU_BrCa_status==1]), 
        population="Cases", model_old="Gail", model_new="Gail + PRS")

# Controls
print("Controls")
nri_tab(table(test_data$gail_10risk_cat[test_data$TEU_BrCa_time > 10], 
              test_data$gail_prs_10risk_adj_cat[test_data$TEU_BrCa_time > 10]), 
        population="Controls", model_old="Gail", model_new="Gail + PRS")

```

# Decision curve analysis

```{r, fig.width=10}

library(dcurves)

name_conv <- list("Treat All" = "Screen all",
                  "Treat None" = "Screen none",
                  "TC_10risk_cal" = "Tyrer-Cuzick",
                  "TC_prs_10risk" = "Tyrer-Cuzick with PRS_BC",
                  "gail_10risk_cal" = "Gail",
                  "gail_prs_10risk" = "Gail with PRS_BC")
name_color <- c("Screen all" = "grey",
                "Screen none" = "black",
                "Tyrer-Cuzick" = "#E69F00",
                "Tyrer-Cuzick with PRS_BC" = "#D55E00",
                "Gail" = "#56B4E9",
                "Gail with PRS_BC" = "#0072B2")

allnone_drop <- function(.data) {
  new_dca <- .data$dca %>% 
    filter(! label %in% c("Treat None")) %>% 
    mutate(label = factor(label))
  .data$dca <- new_dca
  return(.data)
}

nameswap <- function(.data) {
  new_dca <- .data$dca %>% mutate(
    label = factor(label, labels=unlist(name_conv[levels(label)], use.names=FALSE))
    )
  .data$dca <- new_dca
  return(.data)
}

# Net benefit
dca_tc <- dca(TEU_BrCa_status ~ TC_10risk_cal + TC_prs_10risk, 
            test_data, 
            thresholds = seq(0, 0.12, by = 0.001)) %>%
  nameswap() %>%
  plot(smooth = TRUE) +
  labs(title="Tyrer-Cuzick") + 
  theme(legend.position = c(0.6, 0.8)) + 
  scale_color_manual(values = name_color)

dca_tc
ggsave(file.path(config$outputs$figures$interim, "TC_DCA.png"), 
       scale=0.7, width=25, height=15, units="cm")

dca_gail <- dca(TEU_BrCa_status ~ gail_10risk_cal + gail_prs_10risk,
            test_data,
            thresholds = seq(0, 0.12, by = 0.001)) %>%
  nameswap() %>%
  plot(smooth = TRUE) +
  labs(title="Gail") +
  theme(legend.position = c(0.6, 0.8)) + 
  scale_color_manual(values = name_color)

dca_gail
ggsave(file.path(config$outputs$figures$interim, "gail_DCA.png"), 
       scale=0.7, width=25, height=15, units="cm")

# Figure 3
tiff(file.path(config$outputs$figures$final,"Figure3.tiff"), units="in", width=10, height=5, res=300)
gridExtra::grid.arrange(dca_tc, dca_gail, ncol=2)
dev.off()

# Interventions avoided
ia_tc <- dca(TEU_BrCa_status ~ TC_10risk_cal + TC_prs_10risk, 
             data = test_data, 
             thresholds = seq(0, 0.12, by = 0.001),
             as_probability = c("TC_10risk_cal", "TC_prs_10risk")) %>%
  net_intervention_avoided(nper=100) %>%
  allnone_drop() %>%
  nameswap() %>%
  plot(smooth = TRUE) +
  labs(title="Tyrer-Cuzick") + 
  theme(legend.position = c(0.3, 0.8)) + 
  scale_color_manual(values = name_color)

ia_tc
ggsave(file.path(config$outputs$figures$interim, "TC_NIA.png"), 
       scale=0.7, width=25, height=15, units="cm")

ia_gail <- dca(TEU_BrCa_status ~ gail_10risk_cal + gail_prs_10risk,
               test_data,
               thresholds = seq(0, 0.12, by = 0.001)#,
               # as_probability = c("TC_10risk_cal", "TC_prs_10risk")
            ) %>%
  net_intervention_avoided(nper=100) %>%
  allnone_drop() %>%
  nameswap() %>%
  plot(smooth = TRUE) +
  labs(title="Gail") +
  theme(legend.position = c(0.3, 0.8)) + 
  scale_color_manual(values = name_color)

ia_gail
ggsave(file.path(config$outputs$figures$interim, "gail_NIA.png"), 
       scale=0.7, width=25, height=15, units="cm")

gridExtra::grid.arrange(ia_tc, ia_gail, ncol=2)

```

# Supplementary Materials

## Supplementary Figure: Flowchart illustrating selection of analytic dataset (n = `r nrow(data)`).

!["*'Invalid or missing data for core variables' comprises: 1. Missing data for HRT use (Never, Previous or Current) as this variable was required for the Tyrer-Cuzick model. Previous HRT users with unknown age of stopping HRT were categorised into 'Previous (<5)' and 'Previous (>5)' by imputation. 2. Twenty or more sisters, as this crashes the Tyrer-Cuzick model. 3. An age of stopping HRT that was greater than age at baseline."](Figures/Manuscript/Exclusions.png){height=700px}

## Supplementary Table: Baseline characteristics

Including variables from both models, by quintile (1st, 3rd, 5th, overall) of Genomics plcs PRS

```{r, results='asis'}

# Descrp stats
part1 <- tableby(PRS_bc_std_q ~ 
                   # Standard
                   TEU_BaC_AgeCat + Height + Weight +
                   # Reproductive
                   FSF_AgeMenarche + TC_Parity + FSF_AgeMenopause +
                   TC_HRT_imputed +
                   # Diagnoses
                   TC_OvarianCaDx + Biopsy_Prevalent +
                   # Family history
                   TC_MotherBrCa + 
                   TC_NSisters + TC_SiblingBrCa,
                 data=data %>% mutate(Biopsy_Prevalent=factor(Biopsy_Prevalent)), 
                 control=my_controls)

# Descriptive stats within subgroups

# Reproductive
rep_agefb <- tableby(PRS_bc_std_q ~ AgeFirstBirth,
                 data=data %>% filter(!is.na(TC_Parity) & TC_Parity == "Parous"),
                 control=my_controls)

rep_hrtdur <- tableby(PRS_bc_std_q ~ FSF_HRTDuration,
                 data=data %>% filter(TC_HRT_imputed %in% c("Current",
                                                            "Previous user (less than 5 years ago)")),
                 control=my_controls)

rep_hrtsince <- tableby(PRS_bc_std_q ~ FSF_HRT_timesince,
                 data=data %>% filter(TC_HRT_imputed == "Previous user (less than 5 years ago)"),
                 control=my_controls)

# Medical history
dx_ov <- tableby(PRS_bc_std_q ~ TC_OvarianCaAge,
                 data=data %>% filter(TC_OvarianCaDx == "Yes"),
                 control=my_controls)

dx_hyp <- tableby(PRS_bc_std_q ~ Gail_AtypicalHyperplasia,
                  data=data %>% 
                    filter(Biopsy_Prevalent > 0) %>% 
                    mutate(Gail_AtypicalHyperplasia = factor(Gail_AtypicalHyperplasia)),
                  control=my_controls)

# Family history 
fh_ma <- tableby(PRS_bc_std_q ~ TC_MotherAge,
                 data=data %>% filter(TC_MotherBrCa == "No"),
                 control=my_controls)

summary(part1, labelTranslations = pretty_list,
        title=paste0('Table 1. Descriptive statistics by quintiles of breast cancer PRS'))

# Reproductive
summary(rep_agefb, labelTranslations = pretty_list,
        title="Table 1a. Age at first birth among parous women")
summary(rep_hrtdur, labelTranslations = pretty_list,
        title="Table 1b. Duration of HRT use, among women who are currently taking HRT or who previously used HRT less than 5 years ago")
summary(rep_hrtsince, labelTranslations = pretty_list,
        title="Table 1c. Time since stopping HRT among women who had previously used HRT less than 5 years ago")
# Medical history
summary(dx_ov, labelTranslations = pretty_list,
        title="Table 1d. Age at ovarian cancer diagnosis among women diagnosed with ovarian cancer")
summary(dx_hyp, labelTranslations = pretty_list,
        title="Table 1e. Diagnosis of atypical hyperplasia, among women with breast biopsy")
# Family history
summary(fh_ma, labelTranslations = pretty_list,
        title="Table 1f. Maternal age at baseline or death, among women whose mothers had not had breast cancer. Note: Tyrer-Cuzick model asked for maternal age at dx of breast cancer or, if no dx, then current age or age at death. In UKB data, maternal age at breast cancer diagnosis was not available, only age at baseline assessment or age at death.")

```

## Cox regression coefficients of individual variables in TC and Gail models

### TC

```{r, results='asis'}

# Tyrer-Cuzick

tc_varlist <- c("FSF_AgeMenopause", "FSF_AgeMenarche", "Height", "BMI", 
                "AgeFirstBirth", "TC_HRT_imputed", "AtypicalHyperplasia")
                # "TEU_BaC_AgeAtRec", "TC_Parity", 
                # "AtypicalHyperplasia", "TC_OvarianCaDx", "TC_OvarianCaAge", "FSF_HRTDuration", 
                # "FSF_HRT_timesince", "TC_MotherBrCa", "TC_NSisters", "TC_SiblingBrCa")

tc_data <- train_data %>% 
  mutate(BMI = Weight/(Height/100)^2,
         FSF_AgeMenopause = FSF_AgeMenopause/5,
         FSF_AgeMenarche = relevel(
           cut(FSF_AgeMenarche, breaks=c(0, 11, 12, 13, 14, 15, 16, 17, 100),
               labels=c("<11", "11", "12", "13", "14", "15", "16", "17 or older"), right=FALSE),
           ref="13"),
         Height = cut(Height, breaks=c(0, 160, 170, 200),
                      labels=c("<1.6", "1.6-1.7", "1.7 or taller"), right=FALSE),
         BMI = cut(BMI, breaks=c(0, 21, 23, 25, 27, 100),
                   labels=c("<21", "21 to <23", "23 to <25", "25 to <27", "27 or more"), right=FALSE),
         AgeFirstBirth = factor(dplyr::case_when(
           TC_Parity == "Nulliparous" ~ "Nulliparous",
           is.na(AgeFirstBirth) ~ NA,
           AgeFirstBirth < 20 ~ "<17-19",
           AgeFirstBirth < 25 ~ "20-24",
           AgeFirstBirth < 30 ~ "25-29",
           AgeFirstBirth < 35 ~ "30-34",
           AgeFirstBirth >= 35 ~ "35+"
         ), levels=c("Nulliparous", "<17-19", "20-24", "25-29", "30-34", "35+")),
         TC_HRT_imputed = factor(dplyr::case_when(
           TC_HRT_imputed == "Current" ~ "Current",
           TRUE ~ "Not current"
         ), levels=c("Not current", "Current"))) %>%
  select(ID, BaC_Age, TEU_BrCa_time, TEU_BrCa_status, 
                                 all_of(tc_varlist)) %>%
  drop_na

tc_formula <- paste0("Surv(TEU_BrCa_time, TEU_BrCa_status) ~ BaC_Age + ", paste0(tc_varlist, collapse=" + "))
tc_cox <- coxph(as.formula(tc_formula), data=tc_data)

kable(printcoxresults(tc_data, varlist=tc_varlist, modeloutput=tc_cox))

for (var in tc_varlist) {
  tc_formula <- paste0("Surv(TEU_BrCa_time, TEU_BrCa_status) ~ BaC_Age + ", var)
  tc_cox <- coxph(as.formula(tc_formula), data=tc_data)
  print(kable(printcoxresults(tc_data, varlist=var, modeloutput=tc_cox)))
}

```

### Gail

```{r, results='asis'}

#  Gail

gail_varlist <- c("FSF_AgeMenarche", "Biopsy_Prevalent", "AgeCat",
                  "AgeFirstBirth", "FamilyBrCa_HowMany")

gail_data <- train_data %>% 
  mutate(Biopsy_Prevalent = factor(Biopsy_Prevalent),
         FSF_AgeMenarche = relevel(cut(FSF_AgeMenarche, breaks=c(0, 11, 13, 100),
               labels=c("<12", "12-13", ">=14"), right=TRUE), ref=">=14"),
         AgeFirstBirth = factor(dplyr::case_when(
           TC_Parity == "Nulliparous" ~ "25-29 or Nulliparous",
           is.na(AgeFirstBirth) ~ NA,
           AgeFirstBirth < 20 ~ "<20",
           AgeFirstBirth < 25 ~ "20-24",
           AgeFirstBirth < 30 ~ "25-29 or Nulliparous",
           AgeFirstBirth < 35 ~ "30 or older"
         ), levels=c("<20", "20-24", "25-29 or Nulliparous", "30 or older")),
         FamilyBrCa_HowMany = factor(FamilyBrCa_HowMany),
         AgeCat = cut(BaC_Age, breaks=c(0, 50, 100), labels=c("<50yr", ">=50yr"))
         ) %>%
  select(ID, BaC_Age, TEU_BrCa_time, TEU_BrCa_status, 
                                 all_of(gail_varlist)) %>%
  drop_na

gail_cox <- coxph(as.formula(Surv(TEU_BrCa_time, TEU_BrCa_status) ~ FSF_AgeMenarche + 
                   Biopsy_Prevalent + AgeCat + AgeFirstBirth*FamilyBrCa_HowMany), data=gail_data)
print(gail_cox)

gail_cox_nointeract <- coxph(as.formula(Surv(TEU_BrCa_time, TEU_BrCa_status) ~ FSF_AgeMenarche + 
                   Biopsy_Prevalent + AgeCat + AgeFirstBirth + FamilyBrCa_HowMany), data=gail_data)

lmtest::lrtest(gail_cox, gail_cox_nointeract)

for (var in gail_varlist) {
  gail_formula <- paste0("Surv(TEU_BrCa_time, TEU_BrCa_status) ~ BaC_Age + ", var)
  gail_cox <- coxph(as.formula(gail_formula), data=gail_data)
  print(kable(printcoxresults(gail_data, varlist=var, modeloutput=gail_cox)))
}

```

# Distribution of risk

```{r}

# Tyrer-Cuzick
facet_tc <- test_data %>% select(TC_10risk_cal, TC_prs_10risk, TEU_BrCa_status) %>% 
  pivot_longer(cols=c(TC_10risk_cal, TC_prs_10risk))

density_data_tc <- facet_tc %>%
  mutate(
    name = ifelse(TEU_BrCa_status == 0, paste0(name, "_control"), paste0(name, "_case")),
    Model = factor(name, 
                   levels=c("TC_10risk_cal_control", "TC_10risk_cal_case",
                            "TC_prs_10risk_control", "TC_prs_10risk_case"),
                   labels = c("Tyrer-Cuzick controls", "Tyrer-Cuzick cases",
                              "Tyrer-Cuzick with PRS_BC controls", "Tyrer-Cuzick with PRS_BC cases")))

density_TC <- ggplot(density_data_tc, 
       aes(x=value, fill=Model)) + 
  geom_density(alpha=0.4, position="identity") +
  theme(legend.position = c(0.8, 0.8)) +
  xlim(0, 0.2) +
  geom_vline(xintercept=0.05)

# Gail
facet_gail <- test_data %>% select(gail_10risk_cal, gail_prs_10risk, TEU_BrCa_status) %>% 
  pivot_longer(cols=c(gail_10risk_cal, gail_prs_10risk))

density_data_gail <- facet_gail %>%
  mutate(
    name = ifelse(TEU_BrCa_status == 0, paste0(name, "_control"), paste0(name, "_case")),
    Model = factor(name, 
                   levels=c("gail_10risk_cal_control", "gail_10risk_cal_case",
                            "gail_prs_10risk_control", "gail_prs_10risk_case"),
                   labels = c("Gail controls", "Gail cases",
                              "Gail with PRS_BC controls", "Gail with PRS_BC cases")))

density_gail <- ggplot(density_data_gail, 
       aes(x=value, fill=Model)) + 
  geom_density(alpha=0.4, position="identity") +
  theme(legend.position = c(0.8, 0.8)) +
  xlim(0, 0.2) +
  geom_vline(xintercept=0.05)

# Supplementary Figure 5
tiff(file.path(config$outputs$figures$final,"SuppFigureS5.tiff"), units="in", width=8, height=10, res=300)
gridExtra::grid.arrange(density_TC, density_gail, nrow=2)
dev.off()

```

# Fixed risk threshold

```{r}

nri <- nricens(time=test_data$TEU_BrCa_time, event=test_data$TEU_BrCa_status,
        p.std=test_data$TC_10risk_cal, p.new=test_data$TC_prs_10risk,
        t0=10, cut=0.05, msg=FALSE)

# kable(nri$nri)

################
# Tyrer-Cuzick #
################

# Cases
print("Cases")
nri_tab(table(test_data$TC_10risk_cat[test_data$TEU_BrCa_time <= 10 & test_data$TEU_BrCa_status==1], 
              test_data$TC_prs_10risk_cat[test_data$TEU_BrCa_time <= 10 & test_data$TEU_BrCa_status==1]), 
        population="Cases", model_old="Tyrer-Cuzick", model_new="Tyrer-Cuzick + PRS")

# Controls
print("Controls")
nri_tab(table(test_data$TC_10risk_cat[test_data$TEU_BrCa_time > 10], 
              test_data$TC_prs_10risk_cat[test_data$TEU_BrCa_time > 10]), 
        population="Controls", model_old="Tyrer-Cuzick", model_new="Tyrer-Cuzick + PRS")

```

```{r}

nri <- nricens(time=test_data$TEU_BrCa_time, event=test_data$TEU_BrCa_status,
        p.std=test_data$gail_10risk_cal, p.new=test_data$gail_prs_10risk,
        t0=10, cut=0.05, msg=FALSE)

# kable(nri$nri)

################
# Gail #
################

# Cases
print("Cases")
nri_tab(table(test_data$gail_10risk_cat[test_data$TEU_BrCa_time <= 10 & test_data$TEU_BrCa_status==1],
              test_data$gail_prs_10risk_cat[test_data$TEU_BrCa_time <= 10 & test_data$TEU_BrCa_status==1]),
        population="Cases", model_old="Gail", model_new="Gail + PRS")

# Controls
print("Controls")
nri_tab(table(test_data$gail_10risk_cat[test_data$TEU_BrCa_time > 10],
              test_data$gail_prs_10risk_cat[test_data$TEU_BrCa_time > 10]),
        population="Controls", model_old="Gail", model_new="Gail + PRS")

```

# TC v8 with Genomics plc

```{r}

TC_5pct <- table(test_data$TC_10risk_cal > 0.05, useNA='ifany')
TC_5prop <- TC_5pct[2]/nrow(test_data)
TC_Gplc_threshold <- quantile(test_data$TC_10risk_GenomicsPLC_cal, probs=c(1-TC_5prop))

test_data <- test_data %>% mutate(
  TC_Gplc_10risk_adj = TC_10risk_GenomicsPLC_cal - (quantile(TC_10risk_GenomicsPLC_cal, 
                                                                 probs=c(1-TC_5prop)) - 0.05),
  TC_Gplc_10risk_adj_cat = factor(ifelse(TC_10risk_GenomicsPLC_cal <= TC_Gplc_threshold, 
                                          paste0("<=", round(100*TC_Gplc_threshold,2), "%"), 
                                          paste0(">", round(100*TC_Gplc_threshold,2), "%")),
                               levels=c(paste0("<=", round(100*TC_Gplc_threshold,2), "%"), 
                                        paste0(">", round(100*TC_Gplc_threshold,2), "%")))
  )

nri_tc_Gplc <- nricens(time=test_data$TEU_BrCa_time, event=test_data$TEU_BrCa_status,
        p.std=test_data$TC_10risk_cal, p.new=test_data$TC_Gplc_10risk_adj,
        t0=10, cut=0.05, msg=FALSE)

```

```{r}

dt <- data.frame(model=c("Model only", "Model with PRS", "Integrated PRS"),
                 HarrellC=c(
                       HarrellC_CI(df=test_data, score="TC_10risk_cal", 
                              time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2),
                       HarrellC_CI(df=test_data, score="TC_prs_10risk", 
                              time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2),
                       HarrellC_CI(df=test_data, score="TC_10risk_GenomicsPLC_cal", 
                              time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2)),
                 NRI_overall=c(NA, nri_ci(nri_tc, type="overall", dp=3), 
                               nri_ci(nri_tc_Gplc, type="overall", dp=3)),
                 NRI_case=c(NA, nri_ci(nri_tc, type="case", dp=3), 
                            nri_ci(nri_tc_Gplc, type="case", dp=3)),
                 NRI_control=c(NA, nri_ci(nri_tc, type="control", dp=3), 
                               nri_ci(nri_tc_Gplc, type="control", dp=3)))

ft <- flextable(dt) %>%
  # merge_v(j = ~ model + NRI_overall + NRI_case + NRI_control) %>%
  # separate_header(sep="_") %>%
  set_header_labels(model="Tyrer-Cuzick Model", HarrellC="Harrell's C (95% CI)",
                    NRI_overall="Overall", 
                    NRI_case=paste0("Case (N=", nrow(test_data[test_data$TEU_BrCa_time <= 10 &
                                                                 test_data$TEU_BrCa_status==1,]), ")"), 
                    NRI_control=paste0("Control (N=", nrow(test_data[test_data$TEU_BrCa_time > 10,]), ")")) %>%
  add_header_row(values=c("", "NRI"), colwidths=c(2, 3)) %>%
  align(align = "center", part = "all", j=c("HarrellC", 'NRI_overall', 'NRI_case', 'NRI_control')) %>%
  bg(i= ~ is.na(HarrellC), bg="lightgrey", part="body") %>%
  fix_border_issues(part="all") %>%
  set_table_properties(layout="autofit") %>%
  flextable::footnote(i=1, j=3, ref_symbols="a", part="header",
           value=as_paragraph("NRI: Net Reclassification Index for 10yr risks from model with PRS compared to Calibrated model, with a fixed proportion of women classified as high risk")) %>%
  flextable::footnote(i=2, j=3, ref_symbols="b", part="header",
           value=as_paragraph("Overall NRI = Case NRI + Control NRI, where cases are defined as individuals diagnosed with breast cancer within 10 years and controls are defined as individuals who were still at risk of breast cancer by 10 years of follow-up.")) %>%
  flextable::footnote(i=3, j=1, ref_symbols="c", part="body",
           value=as_paragraph("Integrated PRS is the output from the Tyrer-Cuzick model run with the PRS as an input, formatted as a log(OR)"))

ft

```

# Results using Mavaddat 313-SNP PRS

```{r, fig.height=9}

model_TC <- coxph(Surv(TEU_BrCa_time, TEU_BrCa_status) ~
                        TC_10risk_cal + TEU_BrCa_313_PRS_center + GeP_Array + 
                        GeP_PC_1 + GeP_PC_2 + GeP_PC_3 + GeP_PC_4,
                 data=train_data)

test_data$TC_prs313_10risk <- get.risk.coxph_test(model_TC, t0=10, newdata=test_data)

model_gail <- coxph(Surv(TEU_BrCa_time, TEU_BrCa_status) ~
                        gail_10risk_cal + TEU_BrCa_313_PRS_center + GeP_Array + 
                        GeP_PC_1 + GeP_PC_2 + GeP_PC_3 + GeP_PC_4,
                 data=train_data)

test_data$gail_prs313_10risk <- get.risk.coxph_test(model_gail, t0=10, newdata=test_data)

```

```{r}
# How many women are above 5% 10yr risk according to the original model?

TC_5pct <- table(test_data$TC_10risk_cal > 0.05, useNA='ifany')
TC_5prop <- TC_5pct[2]/nrow(test_data)
TC_prs_threshold <- quantile(test_data$TC_prs313_10risk, probs=c(1-TC_5prop))
TC_Mav2019_threshold <- quantile(test_data$TC_10risk_Mavaddat2019_cal, probs=c(1-TC_5prop))

gail_5pct <- table(test_data$gail_10risk_cal > 0.05, useNA='ifany')
gail_5prop <- gail_5pct[2]/nrow(test_data)
gail_prs_threshold <- quantile(test_data$gail_prs313_10risk, probs=c(1-gail_5prop))

# How many women are above 5% 10yr risk after adding the PRS?

TC_combined_5pct <- table(test_data$TC_prs313_10risk > 0.05)

gail_combined_5pct <- table(test_data$gail_prs313_10risk > 0.05)

test_data <- test_data %>% mutate(
  TC_prs313_10risk_adj = TC_prs313_10risk - (quantile(TC_prs313_10risk, probs=c(1-TC_5prop)) - 0.05),
  TC_prs313_10risk_adj_cat = factor(ifelse(TC_prs313_10risk <= TC_prs_threshold, 
                                          paste0("<=", round(100*TC_prs_threshold,2), "%"), 
                                          paste0(">", round(100*TC_prs_threshold,2), "%")),
                               levels=c(paste0("<=", round(100*TC_prs_threshold,2), "%"), 
                                        paste0(">", round(100*TC_prs_threshold,2), "%"))),
  TC_Mav2019_10risk_adj = TC_10risk_Mavaddat2019_cal - (quantile(TC_10risk_Mavaddat2019_cal, 
                                                                 probs=c(1-TC_5prop)) - 0.05),
  TC_Mav2019_10risk_adj_cat = factor(ifelse(TC_10risk_Mavaddat2019_cal <= TC_prs_threshold, 
                                          paste0("<=", round(100*TC_Mav2019_threshold,2), "%"), 
                                          paste0(">", round(100*TC_Mav2019_threshold,2), "%")),
                               levels=c(paste0("<=", round(100*TC_Mav2019_threshold,2), "%"), 
                                        paste0(">", round(100*TC_Mav2019_threshold,2), "%"))),
  gail_prs313_10risk_adj = gail_prs313_10risk - (quantile(gail_prs313_10risk, probs=c(1-gail_5prop)) - 0.05),
  gail_prs313_10risk_adj_cat = factor(ifelse(gail_prs313_10risk <= gail_prs_threshold, 
                                          paste0("<=", round(100*gail_prs_threshold,2), "%"), 
                                          paste0(">", round(100*gail_prs_threshold,2), "%")),
                               levels=c(paste0("<=", round(100*gail_prs_threshold,2), "%"), 
                                        paste0(">", round(100*gail_prs_threshold,2), "%")))
  )

```

```{r, results=FALSE}

# Net reclassification index, for Table 1

# Tyrer-Cuzick
nri_tc <- nricens(time=test_data$TEU_BrCa_time, event=test_data$TEU_BrCa_status,
        p.std=test_data$TC_10risk_cal, p.new=test_data$TC_prs313_10risk_adj,
        t0=10, cut=0.05, msg=FALSE)

nri_tc_mav <- nricens(time=test_data$TEU_BrCa_time, event=test_data$TEU_BrCa_status,
        p.std=test_data$TC_10risk_cal, p.new=test_data$TC_Mav2019_10risk_adj,
        t0=10, cut=0.05, msg=FALSE)

# Gail
nri_gail <- nricens(time=test_data$TEU_BrCa_time, event=test_data$TEU_BrCa_status,
        p.std=test_data$gail_10risk_cal, p.new=test_data$gail_prs313_10risk_adj,
        t0=10, cut=0.05, msg=FALSE)

nri_ci <- function(nri_output, type="overall", dp=3){
  row <- dplyr::case_when(
    type=="overall" ~ 1,
    type=="case" ~ 2,
    type=="control" ~ 3
  )
  est <- nri_output$nri$Estimate[row]
  lci <- nri_output$nri$Lower[row]
  uci <- nri_output$nri$Upper[row]

  return(paste0(pretty_dp(est, dp=dp), " ", pretty_confint(lci, uci, dp=dp)))
}

```

```{r}

dt <- data.frame(model=c("Tyrer-Cuzick", "Model only", "Model with PRS", "Integrated PRS",
                         "Gail", "Model only", "Model with PRS"),
                 HarrellC=c(NA,
                       HarrellC_CI(df=test_data, score="TC_10risk_cal", 
                              time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2),
                       HarrellC_CI(df=test_data, score="TC_prs313_10risk", 
                              time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2),
                       HarrellC_CI(df=test_data, score="TC_10risk_Mavaddat2019_cal", 
                              time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2),
                       NA,
                       HarrellC_CI(df=test_data, score="gail_10risk_cal", 
                              time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2),
                       HarrellC_CI(df=test_data, score="gail_prs313_10risk", 
                              time="TEU_BrCa_time", status = "TEU_BrCa_status", dp=2)),
                 NRI_overall=c(NA, NA, nri_ci(nri_tc, type="overall", dp=3), nri_ci(nri_tc_mav, type="overall", dp=3), 
                               NA, NA, nri_ci(nri_gail, type="overall", dp=3)),
                 NRI_case=c(NA, NA, nri_ci(nri_tc, type="case", dp=3), nri_ci(nri_tc_mav, type="case", dp=3), 
                               NA, NA, nri_ci(nri_gail, type="case", dp=3)),
                 NRI_control=c(NA, NA, nri_ci(nri_tc, type="control", dp=3), nri_ci(nri_tc_mav, type="control", dp=3), 
                               NA, NA, nri_ci(nri_gail, type="control", dp=3)))

ft <- flextable(dt) %>%
  # merge_v(j = ~ model + NRI_overall + NRI_case + NRI_control) %>%
  # separate_header(sep="_") %>%
  set_header_labels(model="Model", HarrellC="Harrell's C (95% CI)",
                    NRI_overall="Overall", 
                    NRI_case=paste0("Case (N=", nrow(test_data[test_data$TEU_BrCa_time <= 10 &
                                                                 test_data$TEU_BrCa_status==1,]), ")"), 
                    NRI_control=paste0("Control (N=", nrow(test_data[test_data$TEU_BrCa_time > 10,]), ")")) %>%
  add_header_row(values=c("", "NRI"), colwidths=c(2, 3)) %>%
  align(align = "center", part = "all", j=c("HarrellC", 'NRI_overall', 'NRI_case', 'NRI_control')) %>%
  bg(i= ~ is.na(HarrellC), bg="lightgrey", part="body") %>%
  merge_at(i=1, j=1:5) %>%
  merge_at(i=5, j=1:5) %>%
  fix_border_issues(part="all") %>%
  set_table_properties(layout="autofit") %>%
  flextable::footnote(i=1, j=3, 
           value=as_paragraph("NRI: Net Reclassification Index for 10yr risks from model with PRS compared to Calibrated model, with a fixed proportion of women classified as high risk"), 
           ref_symbols="a", part="header") %>%
  flextable::footnote(i=2, j=3, 
           value=as_paragraph("Overall NRI = Case NRI + Control NRI, where cases are defined as individuals diagnosed with breast cancer within 10 years and controls are defined as individuals who were still at risk of breast cancer by 10 years of follow-up."), 
           ref_symbols="b", part="header") %>%
  flextable::footnote(i=4, j=1, 
           value=as_paragraph("Integrated PRS is the output from the Tyrer-Cuzick model run with the PRS as an input, formatted as a log(OR)"), 
           ref_symbols="c", part="body")

ft

```